# Antibiotic susceptibility

```{r message=FALSE}
library(tidyverse)
library(usethis)
```

### Vancomycin

Manually curated Lactobacillus database from abx susceptibility paper. Don't take
phenotype info from this paper. The info in table 1 indicates the conditions
under which the strain was grown for the experiment, not all conditions under
which it can be grown. There is more good info on other abx to take from this
paper.

```{r message=FALSE}
susceptibility_lactobacillus <- read_csv( "Lactobacillus_data.csv") %>%
  mutate(antibiotic = "vancomycin") %>%
  select(taxon = name, rank, antibiotic, value = vancomycin, doi) %>%
  mutate(taxon = str_replace_all(taxon, "\\s+", " ")) %>%
  # Policy of this package is not to work with subspecies, but instead to
  # convert subspecies name to an equivalent species name, and pretend the
  # subspecies is just a different species.
  mutate(taxon = str_remove(taxon, " \\w+ subsp\\.")) %>%
  mutate(value = if_else(value, "susceptible", "resistant"))
```

### Other info

```{r message=FALSE}
susceptibility_misc <- read_csv("susceptibility_misc.csv")
```

### Export

```{r}
taxon_susceptibility <- bind_rows(
  susceptibility_misc, susceptibility_lactobacillus)
taxon_susceptibility %>%
  count(antibiotic, value)
```

```{r}
taxon_susceptibility <- as.data.frame(taxon_susceptibility)
usethis::use_data(taxon_susceptibility, overwrite = T)
```

Review paper on drug class:
https://doi.org/10.1016/j.bcp.2017.01.003

Tet protection genes on mobile elements:
http://faculty.washington.edu/marilynr/ https://doi.org/10.1016/j.femsle.2005.02.034

The Clinical and Laboratory Standards Institute (CLSI) guidelines assesses abx
resistance.
